DATA 621 - Business Analytics and Data Mining

### 2020-11-02 HW4/HW4_Draft.Rmd
Update

# Relevel some of the variables to have the default make more sense
# Set default marital status to unmarried
set(DT, NULL, 'MSTATUS', relevel(DT$MSTATUS, 'No'))
# Set default car usage to private
set(DT, NULL, 'CAR_USE', relevel(DT$CAR_USE, 'Private'))
```

@@

 * `OWN` with `HOME_VAL` as discussed in the missing data section
 * `KIDSDRIV` with `HOMEKIDS`
 * A three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`
 * `CLM_FREQ` with `OLDCLAIM`
 * `MSTATUS` with `SEX`
   * One sex may be more affected by marriage than the other

```{r model1DummyFreq}
trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
trnDumFreq <- dummyVars(TARGET_FLAG ~ .  + OWN * HOME_VAL +
                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
                          TARGET_AMT,
                        fullRank = TRUE, data = trnX)
trnXFD <- predict(trnDumFreq, newdata = trnX)
```

@@

```{r model1RunFreq}
trC <- trainControl(method = 'none',            # Using StepAIC
                    classProbs = TRUE,          # Classification
                    summaryFunction = prSummary # Precision/Recall)
)
set.seed(487)
m1FreqFit <- train(x = trnXFD, y = trnFreq, preProcess = c('nzv', 'knnImpute'),
                   trControl = trC, method = 'glmStepAIC',
                   family = binomial(link = 'logit'), direction = 'both',
                   trace = 1)
kable(summary(m1FreqFit)$coefficients, digits = 3L,
      caption = "Model 1 Frequency Regression Output")
```

#### Coefficient Discussion
Positive coefficients represent an **increased** risk of a claim over the
baseline and negative coefficients represent a **decreased** risk over the
baseline. Most of the coefficients are of the expected direction. For example,
having one's license revoked or living in an urban area is a *positive* risk
factor for claims while higher levels of wealth or education are *negative* risk
factors.

Some interesting observations. `SEX` is not a significant predictor, even at the
10% level, but removing it does not increase the AIC, so it remains. The
hypothesized interaction between `MSTATUS` and `SEX` does lower the deviance,
but not enough to justify the additional parameter.

The predictors with the greatest effect on risk are `URBANICITY` and `INCOME`.
Having the three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`,
together with `INCOME` and `BLUEBOOK` alone, obviated the need for stand-alone
`HOME_VAL` and the other two possible interactions.

In terms of variable importance, it can be estimated from the the last trace of
the stepping procedure shown below. The further down the table the variable is,
the more dramatic its inclusion or exclusion will be. The top 10 variables in
order would be: `URBANICITY.Rural`, `REVOKED.Yes`, `CAR_USE.Commercial`,
`TRAVTIME`, `CAR_TYPE.Sports Car`, `CLM_FREQ`, `MVR_PTS`, `TIF`, `CAR_TYPE.SUV`,
and `JOB.Blue Collar`.

```
                             Df Deviance    AIC
<none>                            5164.6 5224.6
+ HOMEKIDS                    1   5162.7 5224.7
+ AGE                         1   5163.1 5225.1
- SEX.F                       1   5167.4 5225.4
+ EDUCATION.Masters           1   5163.4 5225.4
+ JOB.Manager                 1   5163.7 5225.7
+ OLDCLAIM                    1   5164.0 5226.0
+ EDUCATION.PhD               1   5164.0 5226.0
+ `HOME_VAL:BLUEBOOK`         1   5164.2 5226.2
+ CAR_AGE                     1   5164.2 5226.2
+ `MSTATUSYes:SEXF`           1   5164.5 5226.5
+ OWN.Yes                     1   5164.5 5226.5
+ YOJ                         1   5164.6 5226.6
+ RED_CAR.yes                 1   5164.6 5226.6
+ `EDUCATION.High School`     1   5164.6 5226.6
+ `INCOME:HOME_VAL:BLUEBOOK`  1   5164.6 5226.6
- `INCOME:BLUEBOOK`           1   5168.8 5226.8
- `INCOME:HOME_VAL`           1   5170.0 5228.0
- JOB.Lawyer                  1   5170.6 5228.6
- JOB.Student                 1   5171.4 5229.4
- `CAR_TYPE.Panel Truck`      1   5173.5 5231.5
- MSTATUS.Yes                 1   5174.8 5232.8
- CAR_TYPE.Van                1   5179.5 5237.5
- BLUEBOOK                    1   5179.9 5237.9
- `OLDCLAIM:CLM_FREQ`         1   5180.0 5238.0
- `JOB.Home Maker`            1   5180.2 5238.2
- HOME_VAL                    1   5180.7 5238.7
- EDUCATION.Bachelors         1   5181.4 5239.4
- INCOME                      1   5182.3 5240.3
- PARENT1.Yes                 1   5182.6 5240.6
- CAR_TYPE.Pickup             1   5183.2 5241.2
- JOB.Professional            1   5186.1 5244.1
- JOB.Clerical                1   5195.4 5253.4
- KIDSDRIV                    1   5197.8 5255.8
- `JOB.Blue Collar`           1   5201.2 5259.2
- CAR_TYPE.SUV                1   5202.5 5260.5
- TIF                         1   5204.4 5262.4
- MVR_PTS                     1   5204.8 5262.8
- CLM_FREQ                    1   5205.4 5263.4
- `CAR_TYPE.Sports Car`       1   5208.4 5266.4
- TRAVTIME                    1   5211.7 5269.7
- CAR_USE.Commercial          1   5230.2 5288.2
- REVOKED.Yes                 1   5233.0 5291.0
- URBANICITY.Rural            1   5612.1 5670.1
```

@@

#### Training
The observations in the training set which have claim amounts will be extracted
and used to fit the linear regression. Similar to frequency, the model will be
selected via a forward and backward stepwise AIC algorithm starting with every
relevant predictive variable and some selected interactions.

```{r model1DummySev}
trnSev <- trnX$TARGET_AMT[trnX$TARGET_AMT > 0]
trnDumSev <- dummyVars(TARGET_AMT ~ .  + OWN * HOME_VAL +
                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
                          TARGET_FLAG,
                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
trnXSD <- predict(trnDumFreq, newdata = trnX[TARGET_AMT > 0])
```

@@

#### Coefficient Disucssion
Hi!

```
                             Df   Deviance   AIC
<none>                          9.5466e+10 31329
- REVOKED.Yes                 1 9.5614e+10 31329
+ JOB.Manager                 1 9.5368e+10 31330
- MSTATUS.Yes                 1 9.5625e+10 31330
+ CAR_AGE                     1 9.5379e+10 31330
- `OLDCLAIM:CLM_FREQ`         1 9.5645e+10 31330
+ `MSTATUSYes:SEXF`           1 9.5392e+10 31330
- CAR_USE.Commercial          1 9.5650e+10 31330
+ SEX.F                       1 9.5399e+10 31330
+ JOB.Professional            1 9.5400e+10 31330
+ CAR_TYPE.Pickup             1 9.5410e+10 31330
+ CAR_TYPE.Van                1 9.5412e+10 31330
+ MVR_PTS                     1 9.5413e+10 31330
+ HOMEKIDS                    1 9.5415e+10 31330
+ CLM_FREQ                    1 9.5431e+10 31331
- `INCOME:HOME_VAL`           1 9.5688e+10 31331
+ RED_CAR.yes                 1 9.5440e+10 31331
- HOME_VAL                    1 9.5694e+10 31331
+ JOB.Student                 1 9.5444e+10 31331
+ EDUCATION.Masters           1 9.5445e+10 31331
+ PARENT1.Yes                 1 9.5452e+10 31331
- OLDCLAIM                    1 9.5709e+10 31331
+ JOB.Clerical                1 9.5455e+10 31331
+ `CAR_TYPE.Panel Truck`      1 9.5455e+10 31331
+ `EDUCATION.High School`     1 9.5456e+10 31331
+ OWN.Yes                     1 9.5457e+10 31331
+ TRAVTIME                    1 9.5458e+10 31331
+ `INCOME:HOME_VAL:BLUEBOOK`  1 9.5460e+10 31331
+ `JOB.Home Maker`            1 9.5460e+10 31331
+ URBANICITY.Rural            1 9.5463e+10 31331
+ `INCOME:BLUEBOOK`           1 9.5464e+10 31331
+ AGE                         1 9.5464e+10 31331
+ `HOME_VAL:BLUEBOOK`         1 9.5465e+10 31331
+ `CAR_TYPE.Sports Car`       1 9.5465e+10 31331
+ YOJ                         1 9.5465e+10 31331
+ `JOB.Blue Collar`           1 9.5466e+10 31331
+ JOB.Lawyer                  1 9.5466e+10 31331
+ KIDSDRIV                    1 9.5466e+10 31331
+ EDUCATION.Bachelors         1 9.5466e+10 31331
+ CAR_TYPE.SUV                1 9.5466e+10 31331
+ `KIDSDRIV:HOMEKIDS`         1 9.5466e+10 31331
+ TIF                         1 9.5466e+10 31331
+ INCOME                      1 9.5466e+10 31331
+ EDUCATION.PhD               1 9.5466e+10 31331
- BLUEBOOK                    1 9.6124e+10 31337
```
# SELECT MODELS
## Model Selection Criteria
For that *classification* algorithm, we will look at accuracy, AUC, and F1, and
select the model which performs best two out of three in those metrics. For the
*regression* algorithm we will look at RMSE, AIC. and MAE (since the values must
be positive) and select the model which performs best two out of three in those
metrics.

>>>>

# Relevel some of the variables to have the default make more sense
# Set default marital status to unmarried
set(DT, NULL, 'MSTATUS', relevel(DT$MSTATUS, 'No'))
# Set default car usage to private
set(DT, NULL, 'CAR_USE', relevel(DT$CAR_USE, 'Private'))
# Set default are to rural, making urban a positive? risk factor
set(DT, NULL, 'URBANICITY', relevel(DT$URBANICITY, 'Rural'))
```

@@

 * `OWN` with `HOME_VAL` as discussed in the missing data section
 * `KIDSDRIV` with `HOMEKIDS`
 * A three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`
 * `CLM_FREQ` with `OLDCLAIM`
 * `MSTATUS` with `SEX`
   * One sex may be more affected by marriage than the other
 * `CAR_USE` with `URBANICITY`

```{r model1DummyFreq}
trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
trnDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_AMT,
                        fullRank = TRUE, data = trnX)
trnXFD <- predict(trnDumFreq, newdata = trnX)
```

@@

```{r model1RunFreq, cache = 2}
trC <- trainControl(method = 'none',            # Using StepAIC
                    classProbs = TRUE,          # Classification
                    summaryFunction = prSummary # Precision/Recall)
)
set.seed(487)
m1FreqFit <- train(x = trnXFD, y = trnFreq, preProcess = c('nzv', 'knnImpute'),
                   trControl = trC, method = 'glmStepAIC',
                   family = binomial(link = 'logit'), direction = 'both',
                   trace = 0)
kable(summary(m1FreqFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 1 Frequency Regression Output")
```

#### Coefficient Discussion
It is difficult to directly map the predictors to probabilities, as all the
values have been centered and scaled in order to perform the imputation.
However, we can say that positive coefficients represent an **increased** risk
of a claim over the baseline and negative coefficients represent a **decreased**
risk under the baseline. When validating and then testing, the same imputations
will be done and the results will be predicted on the normal probability scale.

Most of the coefficients are of the expected direction. For example, having
one's license revoked, living in an urban area, or increased travel times are
all *positive* risk factors while higher levels of wealth or education are
*negative* risk factors.

Some interesting observations. The predictors with the greatest effect on risk
are `URBANICITY` and `INCOME`. The value of the coefficient for `SEX` implies
that being female decreases the probability of a claim, as suspected, but it is
only significant as a predictor at the 10% level. However, it is kept in the
model as removing it increases the AIC. The hypothesized interaction between
`MSTATUS` and `SEX` does lower the deviance, but not enough to justify the
additional parameter. `BLUEBOOK` value is a negative risk factor. Perhaps people
with more expensive cars drive more carefully?

The full three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK` was
deemed insignificant in light of the existing two-way interactions which are in
the model.

The interaction between `CLM_FREQ` and `OLDCLAIM` did help the model AIC, and
the coefficient indicates that while `OLDCLAIM` did not justify the additional
parameter, its tempering of `CLM_FREQ` **does** justify the additional
parameter.

The interaction between `CAR_USE` and `URBANICITY` is also of interest. It
indicates that commercial use vehicles don't really add to claim frequency in
rural areas---the singleton variable is absent from the model---but they have
an outsized influence on increased claim probabilities in urban areas!

In terms of variable importance, it can be estimated from the the last trace of
the stepping procedure shown below. The further down the table the variable is,
the more dramatic its inclusion or exclusion will be. The top 10 variables in
order would be: `URBANICITY.Urban`, `CAR_USECommercial:URBANICITYUrban`,
`REVOKED.Yes`, `TRAVTIME`, `CAR_TYPE.Sports Car`, `CLM_FREQ`, `TIF`, `MVR_PTS`,
`CAR_TYPE.SUV`, and `JOB.Blue Collar`.

```
                                      Df Deviance    AIC
<none>                                     5152.3 5212.3
+ HOMEKIDS                             1   5150.5 5212.5
+ AGE                                  1   5150.8 5212.8
+ EDUCATION.Masters                    1   5151.0 5213.0
- SEX.F                                1   5155.2 5213.2
+ JOB.Manager                          1   5151.6 5213.6
+ OLDCLAIM                             1   5151.6 5213.6
+ EDUCATION.PhD                        1   5151.7 5213.7
+ `HOME_VAL:BLUEBOOK`                  1   5151.7 5213.7
+ CAR_AGE                              1   5151.8 5213.8
- `INCOME:BLUEBOOK`                    1   5155.9 5213.9
+ OWN.Yes                              1   5152.1 5214.1
+ `MSTATUSYes:SEXF`                    1   5152.2 5214.2
+ RED_CAR.yes                          1   5152.3 5214.3
+ `EDUCATION.High School`              1   5152.3 5214.3
+ YOJ                                  1   5152.3 5214.3
+ `INCOME:HOME_VAL:BLUEBOOK`           1   5152.3 5214.3
+ CAR_USE.Commercial                   1   5152.3 5214.3
- `INCOME:HOME_VAL`                    1   5158.7 5216.7
- JOB.Student                          1   5159.0 5217.0
- JOB.Lawyer                           1   5159.1 5217.1
- `CAR_TYPE.Panel Truck`               1   5160.0 5218.0
- MSTATUS.Yes                          1   5162.6 5220.6
- CAR_TYPE.Van                         1   5166.3 5224.3
- BLUEBOOK                             1   5166.6 5224.6
- `JOB.Home Maker`                     1   5166.7 5224.7
- `OLDCLAIM:CLM_FREQ`                  1   5167.4 5225.4
- EDUCATION.Bachelors                  1   5169.0 5227.0
- HOME_VAL                             1   5169.5 5227.5
- PARENT1.Yes                          1   5169.9 5227.9
- INCOME                               1   5170.3 5228.3
- CAR_TYPE.Pickup                      1   5170.6 5228.6
- JOB.Professional                     1   5174.0 5232.0
- JOB.Clerical                         1   5181.7 5239.7
- KIDSDRIV                             1   5186.1 5244.1
- `JOB.Blue Collar`                    1   5189.2 5247.2
- CAR_TYPE.SUV                         1   5189.8 5247.8
- MVR_PTS                              1   5191.2 5249.2
- TIF                                  1   5191.8 5249.8
- CLM_FREQ                             1   5193.0 5251.0
- `CAR_TYPE.Sports Car`                1   5195.8 5253.8
- TRAVTIME                             1   5198.9 5256.9
- REVOKED.Yes                          1   5220.1 5278.1
- `CAR_USECommercial:URBANICITYUrban`  1   5229.9 5287.9
- URBANICITY.Urban                     1   5422.8 5480.8
```

@@

#### Training
The observations in the training set which have claim amounts greater than 0
will be extracted and used to fit the linear regression. Similar to frequency,
the model will be selected via a forward and backward stepwise AIC algorithm
starting with the same predictors and interactions as in the frequency model.

```{r model1DummySev}
trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
trnDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_FLAG,
                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
```

@@

#### Coefficient Disucssion
It is not surprising that this model has many fewer predictors. Information
which helps in identifying the *existence* of a claim often does not help in
estimating the *magnitude* of the claim.

Similar to the frequency model, it is difficult to directly map the predictors
to claim values, as all the values have been centered and scaled in order to
perform the imputation. However, we can say that positive coefficients represent
**increased** claim sizes over the baseline and negative coefficients represent
**decreased** claim sizes under the baseline. When validating and then testing,
the same imputations will be done and the results will be predicted on the
normal dollar scale.

The predictive variables with the greatest effect on claim size are `OLD_CLAIM`,
`HOME_VAL`, and `BLUEBOOK`. The first and third make the most sense. One who has
had larger claims in the past is likely to have larger claims in the future.
Also, the claim for damages is very strongly correlated with the value of the
vehicle. `HOME_VAL` is a bit strange. Perhaps it's serving as a proxy for
purchased coverage limit. Maybe people with more expensive homes, who buy larger
limits on their homeowners policies, also purchase larger limits on their
vehicular polices.

The interaction between `INCOME` and `HOME_VAL` serves to temper the effect of
`HOME_VAL`. Similarly for the inter

```
                                      Df   Deviance   AIC
<none>                                   9.5456e+10 31329
- REVOKED.Yes                          1 9.5604e+10 31329
+ JOB.Manager                          1 9.5357e+10 31329
- MSTATUS.Yes                          1 9.5620e+10 31330
+ CAR_AGE                              1 9.5370e+10 31330
+ `MSTATUSYes:SEXF`                    1 9.5382e+10 31330
- `OLDCLAIM:CLM_FREQ`                  1 9.5636e+10 31330
- CAR_USE.Commercial                   1 9.5639e+10 31330
+ SEX.F                                1 9.5389e+10 31330
+ JOB.Professional                     1 9.5391e+10 31330
+ CAR_TYPE.Pickup                      1 9.5400e+10 31330
+ MVR_PTS                              1 9.5402e+10 31330
+ CAR_TYPE.Van                         1 9.5403e+10 31330
+ HOMEKIDS                             1 9.5404e+10 31330
+ CLM_FREQ                             1 9.5420e+10 31330
+ `CAR_USECommercial:URBANICITYUrban`  1 9.5426e+10 31330
+ RED_CAR.yes                          1 9.5430e+10 31331
+ JOB.Student                          1 9.5432e+10 31331
- `INCOME:HOME_VAL`                    1 9.5686e+10 31331
+ EDUCATION.Masters                    1 9.5435e+10 31331
- HOME_VAL                             1 9.5695e+10 31331
+ PARENT1.Yes                          1 9.5442e+10 31331
+ JOB.Clerical                         1 9.5445e+10 31331
+ `CAR_TYPE.Panel Truck`               1 9.5446e+10 31331
- OLDCLAIM                             1 9.5700e+10 31331
+ `EDUCATION.High School`              1 9.5446e+10 31331
+ OWN.Yes                              1 9.5447e+10 31331
+ TRAVTIME                             1 9.5448e+10 31331
+ `INCOME:HOME_VAL:BLUEBOOK`           1 9.5450e+10 31331
+ `JOB.Home Maker`                     1 9.5450e+10 31331
+ URBANICITY.Urban                     1 9.5453e+10 31331
+ `INCOME:BLUEBOOK`                    1 9.5454e+10 31331
+ AGE                                  1 9.5454e+10 31331
+ `HOME_VAL:BLUEBOOK`                  1 9.5455e+10 31331
+ `JOB.Blue Collar`                    1 9.5455e+10 31331
+ `CAR_TYPE.Sports Car`                1 9.5455e+10 31331
+ KIDSDRIV                             1 9.5456e+10 31331
+ JOB.Lawyer                           1 9.5456e+10 31331
+ EDUCATION.Bachelors                  1 9.5456e+10 31331
+ YOJ                                  1 9.5456e+10 31331
+ CAR_TYPE.SUV                         1 9.5456e+10 31331
+ INCOME                               1 9.5456e+10 31331
+ `KIDSDRIV:HOMEKIDS`                  1 9.5456e+10 31331
+ TIF                                  1 9.5456e+10 31331
+ EDUCATION.PhD                        1 9.5456e+10 31331
- BLUEBOOK                             1 9.6115e+10 31337
```

# SELECT MODELS
## Model Selection Criteria
For that *classification* algorithm, we will look at accuracy, AUC, and F1, and
select the model which performs best two out of three in those metrics. For the
*regression* algorithm we will look at RMSE, AIC. and MAE (since the values must
be positive) and select the model which performs best two out of three in those
metrics.
